\subsection{The First Refinement}
\label{sec:component_diagrams-tutorial_firstRefinement}


The single component and abstract state machine is now refined into a system comprising two components as shown in Figure 20 below. The first component is the Control Panel and the second the abstract washing machine sub-system. Two connectors enable communication between the two components. The first connector, CI, is used to pass the Washing Program ID (PID) to the washing machine sub-system and the second connector, WMSTATE, passes the status of the sub-system back to the Control Panel to be displayed. The state machine is unchanged except for the addition of a self transition on state IDLE which constrains the sendWaiting operation so that it only sends the waiting status over the WMSTATE connector while the washing machine is idle.
 
Figure 20 - First Refinement of Washing Machine

The external operation, UserStart, in component CP represents the user starting the wash by passing the selected wash program, using a port-send action on connector CI to the washing machine sub-system. The port-send action is shown in Figure 21. Note that the minimum delay of 1 is used. The value pid1 is a parameter representing the non-deterministic sending of any PID.
 
Figure 21 - First Refinement : Port Sends on CI

A corresponding port-wake operation, start, in the washing machine sub-system receives the program ID that will, in a subsequent refinement, be decoded to control the wash. The port-wake guard is shown in Figure 22. 
A further port-wake operation, ignoreStart, manages inadvertent start requests from the user. Note that this is necessary due to a design decision not to constrain the sending of start messages from CP. If WM is not in a state to respond to the start an explicit ignoreStart is needed to avoid the system deadlocking.
 
Figure 22 - First Refinement : Port Wakes on CI

When the washing machine sub-system receives the pid, it responds with a port-send action on connector WMSTATE to inform the Control Panel that the washing machine is now RUNNING as shown in Figure 23
 
Figure 23 - First Refinement : Port Sends on WMSTATE

The Control Panel receives the message from the washing machine sub-system with the port-wake operation Running shown in Figure 24 so that this information can be displayed to the washing machine user.
 
Figure 24 - First Refinement : Port Wakes on WMSTATE

The State Machine Animation facility is now used again to validate this 2-component system and the model checker is run to check for deadlocks as shown in Figure 25. Note that, again, all operations have been covered by the model checker.
 
Figure 25 - ProB Model Checking Coverage for the First Refinement


%%% Local Variables:
%%% mode: latex
%%% TeX-master: "component_diagrams-user_manual"
%%% End:
